---
title: "DATA 609 - Homework 3"
author: "Joshua Sturm"
date: "September 28, 2018"
output:
  pdf_document:
    keep_tex: yes
  github_document:
  html_document:
    highlight: textmate
    theme: sandstone
    code_folding: show
    toc: yes
    toc_float: yes
    smart: yes
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, collapse = T, cache = T)
```

# Chapter 3 Problems

```{r load-packages}
library(ggplot2)
library(dplyr)
library(magrittr)
library(knitr)
```

## 1 (Page 113, exercise #2)
The following table gives the elongation $e$ in inches per inch (in./in.) for a given stress $S$ on a steel wire measured in pounds per square inch (lb/in$^2$). Test the model $e = c_1S$ by plotting the data. Estimate $c_1$ graphically.

```{r}
S <- c(5, seq(from = 10, to = 100, by = 10))
e <- c(0, 19, 57, 94, 134, 173, 216, 256, 297, 343, 390)
sedf <- data.frame(S,e)

# Print table (dataframe)
t(sedf)

# Plot dataframe
ggplot(data = sedf) +
  geom_point(aes(x = S, y = e)) +
  labs(title = "Elongation vs. Stress", x = "Stress", y = "Elongation")
```

We can estimate the slope using the first and 6th points.

\begin{equation*}
c_1 = \frac{173 - 0}{50 - 0} = `r 173/50`
\end{equation*}

## 2 (Page 121, exercise #2.a)
Formulate the mathematical model that minimizes the largest deviation between the data and the line $y = ax + b$. If a computer is available, solve for the estimates of $a$ and $b$.

```{r}
x <- c(1.0, 2.3, 3.7, 4.2, 6.1, 7.0)
y <- c(3.6, 3.0, 3.2, 5.1, 5.3, 6.8)

devdf <- data.frame(x, y)
t(devdf)

# Plot dataframe
ggplot(data = devdf) +
  geom_point(aes(x = x, y = y))
```

We can fit a model using these data points to minimize the deviation.

```{r}
model <- lm(y ~ x, data = devdf)
summary(model)
```

The largest deviation is `r max(abs(model$residuals))`.

The equation for the line of best fit is 
\begin{equation*}
y = `r model$coefficients[2]`x + `r model$coefficients[1]`
\end{equation*}

Viewing the fitted model:

```{r}
ggplot(data = devdf, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm")
```

