---
title: "DATA 609 - Homework 6"
author: "Joshua Sturm"
date: "October 25, 2018"
output:
  pdf_document:
    keep_tex: yes
  github_document:
  html_document:
    highlight: textmate
    theme: sandstone
    code_folding: show
    toc: yes
    toc_float: yes
    smart: yes
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, collapse = T, cache = T)
```

# Chapter 7 Problems

```{r load-packages}
library(ggplot2)
library(dplyr)
library(magrittr)
library(knitr)
```

## 1 (Page 251, exercise #2)
Nutritional Requirements-A rancher has determined that the minimum weekly nutritional requirements for an average-sized horse include 40 lb of protein, 20 lb of carbohydrates, and 45 lb of roughage. These are obtained from the following sources in varying amounts at the prices indicated:

```{r}
c1 <- c("Hay", "Oats", "Feeding blocks", "High-protein concentrate", "Requirements per horse")
Protein <- c(0.5, 1.0, 2.0, 6.0, 40.0)
Carbohydrates <- c(2.0, 4.0, 0.5, 1.0, 20.0)
Roughage <- c(5.0, 2.0, 1.0, 2.5, 45.0)
Cost <- c(1.80, 3.50, 0.40, 1.00, "")

reqdf <- data.frame(c1 = c1, Protein = Protein, Carbohydrates = Carbohydrates, Roughage = Roughage, Cost = Cost)
names(reqdf)[1] <- ""

kable(reqdf)
```

Formulate a mathematical model to determine how to meet the minimum nutritional requirements at minimum cost.

### 1 Solution
We are given four variables:
\begin{align*}
x_1 &= \text{Hay} \\
x_2 &= \text{Oats} \\
x_3 &= \text{Feeding blocks} \\
x_4 &= \text{High-Protein concentrate}
\end{align*}

Our cost function is:
$$
1.80x_1 + 3.5x_2 + 0.4x_3 + 1.0x_4
$$

And we want to minimize it, subject to the constraints:
\begin{align*}
0.5x_1 + x_2 + 2.0x_3 + 6.0x_4 &\leq 40.0 \ \ \text(Protein) \\
2.0x_1 + 4.0x_2 + 0.5x_3 + 1.0x_4 &\leq 20.0 \ \ \text(Carbohydrates) \\
5.0x_1 + 2.0x_2 + 1.0x_3 + 2.5x_4 &\leq 45.0 \ \ \text(Roughage)
\end{align*}

## 2 (Page 284, exercise #1)
For the example problem in this section, determine the sensitivity of the optimal solution to a change in $c_2$ using the objective function $25x_1 + c_2x_2$.

### 2 Solution
Let $Z = 25x_1 + c_2x_2$.

Then $x_1 = \frac{-c_2}{25}$.

The slope of the lumber constraint is $\frac{-2}{3}$, and the slope of the labour constraint is $\frac{-5}{4}$.

Thus, the range of values for which the current extreme point remains optimal is given by the inequality

$$
\frac{-5}{4} \leq \frac{-c_2}{25} \leq \frac{-2}{3}
$$

$$
\frac{50}{3} \leq c_2 \leq \frac{125}{4}
$$



# Resources used:

- http://sites.fas.harvard.edu/~apm121/lectures/lec8-hq.pdf
