---
title: "DATA 609 - Homework 4"
author: "Joshua Sturm"
date: "October 3, 2018"
output:
  pdf_document:
    keep_tex: yes
    template: default.latex
  github_document:
  html_document:
    highlight: textmate
    theme: sandstone
    code_folding: show
    toc: yes
    toc_float: yes
    smart: yes
always_allow_html: yes
header-includes:
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{booktabs}
   - \usepackage{makecell}
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, collapse = T, cache = T)
```

```{r load-packages}
library(ggplot2)
library(dplyr)
library(magrittr)
library(knitr)
library(kableExtra)
library(stringr)
```

# Chapter 5 problems

## 1 (Page 191, exercise #3)
Using Monte Carlo simulation, write an algorithm to calculate an approximation to $\pi$ by considering the number of random points selected inside the quarter circle

\begin{equation*}
Q : x^2 + y^2 = 1, \ \ x \geq 0, \ y \geq 0
\end{equation*}

where the quarter circle is taken to be inside the square

\begin{equation*}
S : 0 \leq x \leq 1 \ \text{and} \ 0 \leq y \leq 1
\end{equation*}

Use the equation $\frac{\pi}{4} = \frac{\text{area}Q}{\text{area}S}$

### 1 Solution
```{r}
options(scipen = 999)
set.seed(76)
# Begin by creating a function to calculate n random points between 0 and 1
pi_montcarl <- function(n){
  
  # User defines n number of x and y points to generate randomly between 0 and 1
  
  x_gen <- runif(n, 0, 1)
  y_gen <- runif(n, 0, 1)
  
  # Given pi = (area Q / area S) * 4, we can plug in the equation for Q
  # Sum(x^2 + y^2 / n * 4)
  
  Q <- x_gen^2 + y_gen^2 <= 1
  
  return(sum(Q) / n * 4)
}

# We can create a vector of sample sizes to build a table
n_vector <- c(10, 50, 100, 500, 1000, 5000, 10000, 50000, 100000)

# Apply function to vector of sample sizes
pi_est <- sapply(n_vector, pi_montcarl)

# Build table of results
kable(cbind(n_vector, pi, pi_est, abs(pi - pi_est)),
      col.names = c('Sample Size', 'Actual $\\pi$', 'Estimated $\\pi$', 'Absolute difference'),
      "latex", booktabs = T, escape = F) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "striped")
```

As the sample size increases, the simulation's accuracy improves.

## 2 (Page 194, exercise #1)
Use the middle-square method to generate
\begin{align*}
&a. \text{ 10 random numbers using } x_0 = 1009.\\
&b. \text{ 20 random numbers using } x_0 = 653217.\\
&c. \text{ 15 random numbers using } x_0 = 3043.\\
&d. \text{ Comment about the results of each sequence. Was there cycling? Did each sequence degenerate rapidly?}
\end{align*}

### 2 Solution
```{r}
# Function to generate random numbers using the middle-square method
msq <- function(seed, n){
  
  # Function takes two arguments - n iterations and a starting seed number
  
  # initialize vector with seed value
  val_list <- vector("integer", n)
  val_list[1] <- seed
  
  i <- 2
  while(i <= n){
    
    # If seed^2 has an odd number of digits, add a leading 0
    # if((str_length(seed^2) %% 2 != 0)){
    #   seedsq <- paste0(0, seed^2)
    # } else{
    #   seedsq <- seed^2
    # }
    
    # I realized my mistake with the above code - there always needs to be 2*seed digits
    # I can use the str_pad() function to add as many zeros as I want
    seedsq <- as.character(str_pad(string = as.character(seed^2), width = 2*str_length(val_list[1]), side = 'left', pad = '0'))
    
    midseedsq <- as.character(str_sub(seedsq, start = ((str_length(seedsq) / 2) - (str_length(seed) / 2) + 1),  end = (str_length(seedsq) / 2) + (str_length(seed) / 2)))
    # midseedsq <- str_sub(seedsq, start = str_length(seed) - (str_length(seed) / 2) + 1, end = str_length(seed) + (str_length(seed) / 2))
    
    val_list[i] <- as.integer(midseedsq)
    
    seed <- as.integer(midseedsq)
    i <- i + 1
  }
  return(val_list)
}
```

